noam_data_exploration_v1

### 2024-08-21 noam_data_exploration_v3.Rmd
Continued troubleshooting issues with log transformed slope. May drop this issue in general. Continued updating models and figures to be at island_side instead of site level, added more effects-style plots.

  # fit gls for current time series
  loop_mod <- gls(richness ~ year, correlation =corAR1(form=~year), data = newData)

@@

slope_mod <- glmmTMB(Slope ~ habitat*site + (1|site/transect), data = gls_coefficients_merged)
# model convergence problem
summary(slope_mod) # pretty low variance in random effects
plot(gls_res <- simulateResiduals(slope_mod)) # there are issues, but I can't figure out what they are...
# I think this is due to 1200 obs in model
hist(residuals(slope_mod))
plot(predict(slope_mod) ~ residuals(slope_mod))
Anova(slope_mod)
# Response: Slope
#                 Chisq Df Pr(>Chisq)    
# habitat      1397.999  3  < 2.2e-16 ***
# site           39.095  5  2.273e-07 ***
# habitat:site 1738.155 15  < 2.2e-16 ***
r.squaredGLMM(slope_mod) # 0.22, 0.24
plot(allEffects(slope_mod))
# backreef: richness increases on east and north shore, but relatively constant west side
# Fringe: richness decreasing on east and west side, but relatvely constant on north shore
# outer_10: richness neutral/decreasing on east side, but neutral/increasing on north and west side
# outer_17: richness decreasing north/east side, but neutral west side
# Worth noting trends are slight - arguable whether they are biologically meaningful

slope_emm <- emmeans(slope_mod, ~ habitat*site) # create emmeans object
emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
# Note: not sure I want to include this
emmeans(slope_mod, pairwise ~ site | habitat)

@@

```{r glmm approach, fig.height = 20, fig.width=15}
alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
# Framing - is adiv generally higher in certain habitats/regions?

@@

summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
Anova(richness_mod_2)
plot(allEffects(richness_mod_2))
mod_emm <- emmeans(richness_mod_2, ~ habitat * island_side * year_factor) # if we want to compare habitats year to year
richness_plot_info <- plot(mod_emm, plotit = FALSE)
(emmeans_graph <- 
  ggplot(data =richness_plot_info) +
  geom_point(aes(x = exp(the.emmean), y = year_factor, colour = island_side), size = 8) +
  geom_errorbar(aes(xmin = exp(asymp.LCL), xmax = exp(asymp.UCL), y = year_factor, colour = island_side), width = 0.2, size = 1.5) +
  facet_wrap(~habitat) +
  theme_classic() +
  labs(y = "Year\n", x = "Estimated marginal mean richness") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )

@@

```{r total turnover, fig.height = 24, fig.width=16}
# the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
total_turnover<-
  turnover(
  df = long_data_sub,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "location",
  metric = "total"
)

@@

# turnover bounded between 0 and 1 --> beta regression
# need to transform due to presence of 0s and 1s
# Smithson verkuilen transformation for beta-regression
sv_trans = function(prop, N, s = 0.000005){
  (prop*(nrow(N) - 1) + s)/nrow(N)
  # where prop is the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}
total_turnover$Turnover_trans <- sv_trans(total_turnover$total, total_turnover)

# model
total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*island_side +
                                 (1 | island_side/site/transect) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = total_turnover)
summary(total_turnover_mod1) # low temporal autocorrelation... maybe don't need ar1 term?
plot(total_turnover_res <- simulateResiduals(total_turnover_mod1)) # residuals look very weird
Anova(total_turnover_mod1) # obviously everything sig

# total_turnover_mod2 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site + (1 | site/transect/location),
#                                family = beta_family(), data = total_turnover)
# summary(total_turnover_mod2) # 
# plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
# Anova(total_turnover_mod2) 
# anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
# AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
# # need to figure out residuals problem
# plot(allEffects(total_turnover_mod1))
#

# plot
total_turnover_means <-
  total_turnover %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat) %>%
  dplyr::summarise(Turnover_mean = mean(total), 
            Turnover_se = sd(total)/sqrt(n()))

(turnover_plot <-
  ggplot(data = total_turnover_means) +
  geom_point(aes(x = year, y = Turnover_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = year, y = Turnover_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = island_side,
                    group = island_side_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = island_side_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~island_side_habitat, nrow = 6)+
  labs(title = "a. Turnover\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#
```

```{r appearances, fig.height = 24, fig.width=16}
# examine appearance and disappearances independently 
# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points

@@

# model
appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site +
                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = appearance)
summary(appearance_turnover_mod1) # low temporal autocorrelation and it's negative
plot(appearance_turnover_res <- simulateResiduals(appearance_turnover_mod1)) # similar residuals problem
Anova(appearance_turnover_mod1) # obviously everything sig

appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
                               family = beta_family(), data = appearance)
summary(appearance_turnover_mod2) # 
plot(appearance_turnover_res2 <- simulateResiduals(appearance_turnover_mod2)) #
Anova(appearance_turnover_mod2) 
anova(appearance_turnover_mod2, appearance_turnover_mod1) # models are sig different from each other
AICc(appearance_turnover_mod2, appearance_turnover_mod1) # first model is better but not as drastic as total

@@

appearance_means <-
  appearance %>%
  dplyr::group_by(year, site, habitat, site_habitat) %>%
  dplyr::summarise(Appearance_mean = mean(appearance), 
            Appearance_se = sd(appearance)/sqrt(n()))

(appearance_plot <-
  ggplot(data = appearance_means) +
  geom_point(aes(x = year, y = Appearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = year, y = Appearance_mean, colour = site, group = site_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = site,
                    group = site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site_habitat, nrow = 6)+
  labs(title = "b. Appearances\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#
```

```{r disappearances, fig.height = 24, fig.width=16}
disappearance <-
  turnover(
  df = long_data_sub,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "location",
  metric = "disappearance"
)

disappearance <- merge(turnover_meta, disappearance, by = c("year", "location"))
disappearance$Disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
disappearance$year_factor <- as.factor(disappearance$year)

# model
disappearance_turnover_mod1 <- glmmTMB(Disappearance_trans ~ year_factor*habitat*site +
                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = disappearance)
summary(disappearance_turnover_mod1) # strong autocorrelation here
plot(disappearance_turnover_res <- simulateResiduals(disappearance_turnover_mod1)) # 
Anova(disappearance_turnover_mod1) #

disappearance_turnover_mod2 <- glmmTMB(Disappearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
                               family = beta_family(), data = disappearance)
summary(disappearance_turnover_mod2) # 
plot(disappearance_turnover_res2 <- simulateResiduals(disappearance_turnover_mod2)) #
Anova(disappearance_turnover_mod2) 
anova(disappearance_turnover_mod2, disappearance_turnover_mod1) # only case where models are not sig different from each other
AICc(disappearance_turnover_mod2, disappearance_turnover_mod1) # models almost equivalent

@@

disappearance_means <-
  disappearance %>%
  dplyr::group_by(year, site, habitat, site_habitat) %>%
  dplyr::summarise(Disappearance_mean = mean(disappearance), 
            Disappearance_se = sd(disappearance)/sqrt(n()))

(disappearance_plot <-
  ggplot(data = disappearance_means) +
  geom_point(aes(x = year, y = Disappearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = year, y = Disappearance_mean, colour = site, group = site_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = site,
                    group = site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site_habitat, nrow = 6)+
  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
```

>>>>

  # fit gls for current time series
  loop_mod <- gls(logRich ~ year, correlation =corAR1(form=~year), 
                  control=glsControl(msVerbose=T), data = newData)

@@

# slope_mod <- glmmTMB(Slope ~ habitat*site + (1|site/transect), data = gls_coefficients_merged)
# # model convergence problem
# summary(slope_mod) # pretty low variance in random effects
# plot(gls_res <- simulateResiduals(slope_mod)) # there are issues, but I can't figure out what they are...
# # I think this is due to 1200 obs in model
# hist(residuals(slope_mod))
# plot(predict(slope_mod) ~ residuals(slope_mod))
# Anova(slope_mod)
# # Response: Slope
# #                 Chisq Df Pr(>Chisq)    
# # habitat      1397.999  3  < 2.2e-16 ***
# # site           39.095  5  2.273e-07 ***
# # habitat:site 1738.155 15  < 2.2e-16 ***
# r.squaredGLMM(slope_mod) # 0.22, 0.24
# plot(allEffects(slope_mod))
# # backreef: richness increases on east and north shore, but relatively constant west side
# # Fringe: richness decreasing on east and west side, but relatvely constant on north shore
# # outer_10: richness neutral/decreasing on east side, but neutral/increasing on north and west side
# # outer_17: richness decreasing north/east side, but neutral west side
# # Worth noting trends are slight - arguable whether they are biologically meaningful
# 
# 
# 
# 
# slope_emm <- emmeans(slope_mod, ~ habitat*site) # create emmeans object
# emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
# # Note: not sure I want to include this
# emmeans(slope_mod, pairwise ~ site | habitat)
#

@@

```{r glmm approach, fig.height = 15, fig.width=25}
alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
# Framing - is adiv generally higher in certain habitats/regions?

@@

summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
Anova(richness_mod_2)
plot(allEffects(richness_mod_2))
mod_emm <- emmeans(richness_mod_2, ~ habitat * island_side * year_factor) # if we want to compare habitats year to year
richness_plot_info <- plot(mod_emm, plotit = FALSE)
richness_plot_info$year_numeric <- as.numeric(levels(richness_plot_info$year_factor))
(emmeans_graph <- 
  ggplot(data =richness_plot_info) +
  geom_point(aes(y = exp(the.emmean), x = year_numeric, colour = island_side), size = 8) +
  geom_errorbar(aes(ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), x = year_numeric, colour = island_side), 
                width = 0.2, size = 1.5) +
  geom_segment(aes(y = 5.2, yend = 5.2, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
  facet_wrap(~habitat) +
  theme_classic() +
  labs(x = "Year", y = "Estimated marginal mean richness\n") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )

@@

```{r total turnover, fig.height = 16, fig.width=24}
# the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
total_turnover<-
  turnover(
  df = long_data_sub,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "location",
  metric = "total"
)

@@

# turnover bounded between 0 and 1 --> beta regression
# need to transform due to presence of 0s and 1s
# Smithson verkuilen transformation for beta-regression
sv_trans = function(prop, N, s = 0.000005){
  (prop*(nrow(N) - 1) + s)/nrow(N)
  # where prop is the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}

# for backtransforming logit transformed data in graph
inv_logit <- function(x){
  exp(x)/(1+exp(x))
}

total_turnover$Turnover_trans <- sv_trans(total_turnover$total, total_turnover)

# model
total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*island_side +
                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = total_turnover) # convergence problem, probably overspecified
summary(total_turnover_mod1) # low temporal autocorrelation... maybe don't need ar1 term?
plot(total_turnover_res <- simulateResiduals(total_turnover_mod1)) # residuals look very weird
Anova(total_turnover_mod1) # obviously everything sig
total_turnover_emm <- emmeans(total_turnover_mod1, ~ year_factor*habitat*island_side)
total_turnover_plot_info <- plot(total_turnover_emm, plotit = F)
total_turnover_plot_info$year_numeric <- as.numeric(levels(total_turnover_plot_info$year_factor))

(total_turnover_emmeans_graph <- 
  ggplot(data =total_turnover_plot_info) +
  geom_point(aes(y = inv_logit(the.emmean), x = year_numeric, colour = island_side), size = 8) +
  geom_errorbar(aes(ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), x = year_numeric, colour = island_side), 
                width = 0.2, size = 1.5) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
  facet_wrap(~habitat) +
  theme_classic() +
  labs(x = "Year", y = "Estimated marginal mean total turnover\n") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )

# plot
#total_turnover_means <-
#  total_turnover %>%
#  dplyr::group_by(year, island_side, site, habitat, site_habitat) %>%
#  dplyr::summarise(Turnover_mean = mean(total), 
#            Turnover_se = sd(total)/sqrt(n()))
#
#(turnover_plot <-
#  ggplot(data = total_turnover_means) +
#  geom_point(aes(x = year, y = Turnover_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8, stroke = #2) +
#  geom_line(aes(x = year, y = Turnover_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
#  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = island_side,
#                    group = island_side_habitat), width = 0.5, size = 1.2) +
#  scale_colour_manual(values = island_side_colours) +
#  scale_shape_manual(values = habitat_shapes) +
#  facet_wrap(~island_side_habitat, nrow = 6)+
#  labs(title = "a. Turnover\n", y = "Total turnover\n") +
#  theme_bw() + 
#  theme(panel.grid.major = element_blank(), 
#        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
#        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
#        axis.line.y = element_line(colour = "black"),
#        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
#        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
#        axis.line.x = element_line(colour = "black"),
#        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
#        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
#        strip.text.x = element_text(size = 32),
#        legend.key.size = unit(2, "cm"), # increase legend size
#        legend.text = element_text(size = rel(2)), # increase legend text
#        legend.title = element_text(size = rel(2.25)),
#        legend.position = 'bottom', # move legend to bottom
#        legend.box = 'horizontal',
#        legend.box.background = element_rect(colour = "black") # make legend horizontal
#        )
#    )
#
```

```{r appearances, fig.height = 16, fig.width=24}
# examine appearance and disappearances independently 
# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points

@@

# model
appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ year_factor*habitat*island_side +
                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = appearance) # overspecified
summary(appearance_turnover_mod1) # low temporal autocorrelation and it's negative
plot(appearance_turnover_res <- simulateResiduals(appearance_turnover_mod1)) # similar residuals problem
Anova(appearance_turnover_mod1) # obviously everything sig

# appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
#                                family = beta_family(), data = appearance)
# summary(appearance_turnover_mod2) # 
# plot(appearance_turnover_res2 <- simulateResiduals(appearance_turnover_mod2)) #
# Anova(appearance_turnover_mod2) 
# anova(appearance_turnover_mod2, appearance_turnover_mod1) # models are sig different from each other
# AICc(appearance_turnover_mod2, appearance_turnover_mod1) # first model is better but not as drastic as total

appearance_turnover_emm <- emmeans(appearance_turnover_mod1, ~ year_factor*habitat*island_side)
appearance_turnover_plot_info <- plot(appearance_turnover_emm, plotit = F)
appearance_turnover_plot_info$year_numeric <- as.numeric(levels(appearance_turnover_plot_info$year_factor))

(appearance_turnover_emmeans_graph <- 
  ggplot(data =appearance_turnover_plot_info) +
  geom_point(aes(y = inv_logit(the.emmean), x = year_numeric, colour = island_side), size = 8) +
  geom_errorbar(aes(ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), x = year_numeric, colour = island_side), 
                width = 0.2, size = 1.5) +
  geom_segment(aes(y = 0.5, yend = 0.5, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
  facet_wrap(~habitat) +
  theme_classic() +
  labs(x = "Year", y = "Estimated marginal mean appearance turnover\n") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )

@@

# appearance_means <-
#   appearance %>%
#   dplyr::group_by(year, site, habitat, site_habitat) %>%
#   dplyr::summarise(Appearance_mean = mean(appearance), 
#             Appearance_se = sd(appearance)/sqrt(n()))
# 
# (appearance_plot <-
#   ggplot(data = appearance_means) +
#   geom_point(aes(x = year, y = Appearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
#   geom_line(aes(x = year, y = Appearance_mean, colour = site, group = site_habitat), size = 1.2) +
#   geom_errorbar(aes(x = year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = site,
#                     group = site_habitat), width = 0.5, size = 1.2) +
#   scale_colour_manual(values = site_colours) +
#   scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~site_habitat, nrow = 6)+
#   labs(title = "b. Appearances\n", y = "Total turnover\n") +
#   theme_bw() + 
#   theme(panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         panel.background = element_blank(),
#         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
#         axis.line.y = element_line(colour = "black"),
#         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
#         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
#         axis.line.x = element_line(colour = "black"),
#         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
#         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
#         strip.text.x = element_text(size = 32),
#         legend.key.size = unit(2, "cm"), # increase legend size
#         legend.text = element_text(size = rel(2)), # increase legend text
#         legend.title = element_text(size = rel(2.25)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal',
#         legend.box.background = element_rect(colour = "black") # make legend horizontal
#         )
#     )
# #
```

```{r disappearances, fig.width = 24, fig.height=16}
disappearance <-
  turnover(
  df = long_data_sub,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "location",
  metric = "disappearance"
)

disappearance <- merge(turnover_meta, disappearance, by = c("year", "location"))
disappearance$disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
disappearance$year_factor <- as.factor(disappearance$year)

# model
disappearance_turnover_mod1 <- glmmTMB(disappearance_trans ~ year_factor*habitat*island_side +
                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                               family = beta_family(), data = disappearance) # overspecified
summary(disappearance_turnover_mod1) # low temporal autocorrelation and it's negative
plot(disappearance_turnover_res <- simulateResiduals(disappearance_turnover_mod1)) # similar residuals problem
Anova(disappearance_turnover_mod1) # obviously everything sig

# disappearance_turnover_mod2 <- glmmTMB(disappearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
#                                family = beta_family(), data = disappearance)
# summary(disappearance_turnover_mod2) # 
# plot(disappearance_turnover_res2 <- simulateResiduals(disappearance_turnover_mod2)) #
# Anova(disappearance_turnover_mod2) 
# anova(disappearance_turnover_mod2, disappearance_turnover_mod1) # models are sig different from each other
# AICc(disappearance_turnover_mod2, disappearance_turnover_mod1) # first model is better but not as drastic as total

disappearance_turnover_emm <- emmeans(disappearance_turnover_mod1, ~ year_factor*habitat*island_side)
disappearance_turnover_plot_info <- plot(disappearance_turnover_emm, plotit = F)
disappearance_turnover_plot_info$year_numeric <- as.numeric(levels(disappearance_turnover_plot_info$year_factor))

(disappearance_turnover_emmeans_graph <- 
  ggplot(data =disappearance_turnover_plot_info) +
  geom_point(aes(y = inv_logit(the.emmean), x = year_numeric, colour = island_side), size = 8) +
  geom_errorbar(aes(ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), x = year_numeric, colour = island_side), 
                width = 0.2, size = 1.5) +
  geom_segment(aes(y = 0.5, yend = 0.5, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
  facet_wrap(~habitat) +
  theme_classic() +
  labs(x = "Year", y = "Estimated marginal mean disappearance turnover\n") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )

@@

# disappearance_means <-
#   disappearance %>%
#   dplyr::group_by(year, site, habitat, site_habitat) %>%
#   dplyr::summarise(Disappearance_mean = mean(disappearance), 
#             Disappearance_se = sd(disappearance)/sqrt(n()))
# 
# (disappearance_plot <-
#   ggplot(data = disappearance_means) +
#   geom_point(aes(x = year, y = Disappearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
#   geom_line(aes(x = year, y = Disappearance_mean, colour = site, group = site_habitat), size = 1.2) +
#   geom_errorbar(aes(x = year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = # site,
#                     group = site_habitat), width = 0.5, size = 1.2) +
#   scale_colour_manual(values = site_colours) +
#   scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~site_habitat, nrow = 6)+
#   labs(title = "c. Disappearance\n", y = "Total turnover\n") +
#   theme_bw() + 
#   theme(panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         panel.background = element_blank(),
#         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
#         axis.line.y = element_line(colour = "black"),
#         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
#         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
#         axis.line.x = element_line(colour = "black"),
#         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
#         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
#         strip.text.x = element_text(size = 32),
#         legend.key.size = unit(2, "cm"), # increase legend size
#         legend.text = element_text(size = rel(2)), # increase legend text
#         legend.title = element_text(size = rel(2.25)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal',
#         legend.box.background = element_rect(colour = "black") # make legend horizontal
#         )
#     )
```
